---
title: "A (Hopefully Gentle) Introduction to S7"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BigNum}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(S7)
```

This vignette is written to get you acquainted with `{BigNum}` and S7, and assumes no knowledge of the latter; however, this does assume basic general knowledge of objected oriented principles and terminology. I highly recommend that you use this package in conjunction with the [S7 website](https://rconsortium.github.io/S7/), but ideally you should be able to gain a basic understanding of S7 solely through this vignette. If you are completely new to S7, start with this vignette, check the website as you go (especially if I don't explain something to your liking), and once you've read a section or understood a concept, check the source code of `{BigNum}` to see some (ideally less trivial) examples.

It may be paced too slowly, in which case feel free to skip portions you find trivial. This package and the S7 website tend to have more fleshed out examples and more details.
<!-- We will primarily be using examples from the package. -->

## Classes, Properties, & Objects

[Website](https://rconsortium.github.io/S7/articles/classes-objects.html)

We will start with the example from the readme--`Shape`s. We will be constructing a type hierarchy with squares, rectangles, circles, and shapes, with the typical relationships (e.g. all squares are rectangles, all rectangles and circles are shapes). Let us start with `Shape`s though.

```{r}
Shape <- new_class("Shape")
```

This defines a `Shape` class which inherits from the base S7 object and doesn't have any data associated with it. The first argument to `new_class()` as specified in the docs (the S7 website) should be the name of the class as a string. Note that we have only defined a *constructor* and don't have an instance of it.

```{r}
Shape # class
```

To create *a* `Shape`, we can call the constructor. Since we haven't given any data to `Shape`, the default constructor doesn't take any arguments.

```{r}
Shape() # object
```

Let us add some data. Let's say that all `Shape`s should store their name, as a character vector.

```{r}
Shape <- new_class("Shape", properties = list(name = class_character))
```

::: {.callout-note collapse="true"}
## Subjective Style Note
This note can be ignored as it pertains only to style. 

I think that class definitions in general shouldn't be inlined like they are above--however, since it is so short, I think this is an ok exception. Later we will see longer class definitions and will comment on style again there.
:::

We use `properties` in `new_class()` to define what data the class holds. S7 objects have *properties* (similar-ish to S4 `slots`), and are defined through a list. We provide the name of the property: `name`, and the type of data it can hold: `class_character`. `class_character` is part of a special set of properties exported by `{S7}`; other common ones include `class_numeric`, `class_logical`, `class_list`, etc. You can find an exhaustive list [on the website](https://rconsortium.github.io/S7/reference/index.html#compatibility). Defining the class of properties upfront allows for some validation through type safety.

We access properties using the `@` syntax, e.g. `object@property`.

```{r}
square <- Shape("square")
square@name
```

It is important to define the correct property type as R will now assert that the property matches the defined type. 

```{r}
#| error: true
Shape(10)
```

Here we attempt to make a `Shape` with a `@name`[^1] of `10`, but since `10` isn't a character, we cannot do this. Note that there is no automatic type coercion. By default, we pass properties to the constructor by position, and the order is determined by the order in `properties`. We can observe properties of the class `Shape` itself the same way we look at properties of S7 objects.

[^1]: I use the `@` symbol (`@property`) to signify that I am talking about a property of some object, just like how you would access it. This is another style thing, and is not standard.

```{r}
Shape@constructor
```

That is the default constructor--we'll learn how to write a custom constructor later. Let's look at what happens when we add another property.

```{r}
Shape <- new_class("Shape",
  properties = list(
    name = class_character,
    sides = class_integer
  )
)
```

::: {.callout-note collapse="true"}
## Subjective Style Note
Note that this class definition is no longer inlined. In general, I think that (excepting the `name` argument of `new_class()`) one should space arguments out with linebreaks. Here is a full class with the styling I think one should use.

```{r}
#| eval: false
example <- new_class("example",
  parent = parent_class,
  package = "my_package",
  properties = list(
    prop1 = new_property(class_any,
      getter = function(self) {
        self@prop1
      },
      setter = function(self, value) {
        self
      }
    ),
    prop2 = class_any,
    # ...
  ),
  constructor = function(prop1, prop2 = NULL) {
    force(prop1)
    force(prop2)
    new_object(parent_class, prop1 = prop1, prop2 = prop2)
  },
  validator = function(self) {
    if (self@prop2 == TRUE) {
      "@prop2 must be FALSE"
    } else if (self@prop1 != 42) {
      "@prop1 must be 42"
    }
  }
)
```
:::

Properties have default values. We can observe them by creating an empty `Shape`:

```{r}
Shape()
```

or by looking at the default values in the constructor:

```{r}
Shape@constructor
```

Let's add some more `Shape`s, starting with a `Circle`. All circles are shapes, so we wish to reflect that in our class definition. We can do that using inheritance, which in S7 is expressed as follows:

```{r}
Circle <- new_class("Circle",
  parent = Shape,
  properties = list(
    radius = class_numeric
  )
)
```

`new_class()` has a `parent` argument which defaults to `S7_object` which is the parent class of all S7 objects (think `Object` in Java.) We can see this using `class()`[^2]

[^2]: S7 is [built atop of S3](https://rconsortium.github.io/S7/articles/compatibility.html), so methods for S3 OOP will still work on S7 (e.g. `{sloop}` and `class()`). `S7` sometimes exposes some functions which you should use instead, though. Check if there is a `S7` function before defaulting to a S3 one. (e.g. use `S7_class()` instead of checking if `"S7_object"` is in `class(your_object)`).

```{r}
class(Shape())

class(Circle())
```

Inheritance works as you might expect: `Circle`s now have all the methods of `Shape`s, as well as `@radius`.

Think about if the following would return an error, and why. (Hint: check the class definition for `Shape` again.)

```{r}
#| eval: false
Shape <- new_class("Shape",
  properties = list(
    name = class_character,
    sides = class_integer
  )
)

Circle("circle", 1, 5)
```

We defined `@sides` to be an integer, and there is no type coercion, so `1`, a double, won't be accepted.

```{r}
#| error: true
Circle("circle", 1, 5)
```

So we must pass an integer number of sides.

```{r}
my_circle <- Circle("circle", 1L, 5)
my_circle
```

Observe that the default constructor places the properties of the parent class before those of the children.

```{r}
Circle@constructor
```

You can change this easily using a custom constructor, which we will see later.

Whilst we have some validation to ensure our properties are sensible since `@radius` and `@sides` have to be integers and numerics respectively, we can still pass negative values.

```{r}
Circle("circle", -1L, -10)
```

This is obviously not desirable. We can restrict this behavior by creating a custom property.

```{r}
positive_numeric <- new_property(class_numeric,
  validator = function(value) {
    if (value <= 0) "must be greater than 0"
  }
)
```

`validator()` is how a property's values get validated. The first argument we are passing to `new_property()` is `class_numeric`, telling us that data defined by this property can only be of the class `numeric`. Additionally, our validator checks values, and returns a string if there is an issue. Read more about how validation works [here](https://rconsortium.github.io/S7/articles/classes-objects.html#validation-1) but for our purposes, we can just think of a validator as accepting or rejecting values; returning a string rejects a value. `positive_numeric` is defined to only accept positive numerics, and provides an appropriate error message if faced with negative numbers.

```{r}
Shape <- new_class("Shape",
  properties = list(
    name = class_character,
    sides = new_property(class_integer,
      validator = function(value) {
        if (value <= 0) "must be greater than 0"
      }
    )
  )
)

Circle <- new_class("Circle",
  parent = Shape,
  properties = list(
    radius = positive_numeric
  )
)
```


Before we see how our new classes have appropriate and intuitive behavior, note that I inlined a definition of a custom property analogous to `positive_numeric` but restricted to `integer`s, for `@sides`. I suggest only doing this when you don't intend to reuse the property--it adds some clutter but defining `positive_integer` and never using it seems more wasteful. Maybe that is better still, though--I'm torn.

```{r}
#| error: true

Shape("myshape", 0L)

Circle("circle", -1L, -10)

Circle("circle", 1L, -10)
```

Note that the property name (e.g. `@radiues` or `@sides`) is automatically added to the validator's error message. Don't hardcode that yourself. 

Let us further explore properties.

```{r}
positive_numeric <- new_property(class_numeric,
  validator = function(value) {
    if (value <= 0) "must be greater than 0"
  }
)

Shape <- new_class("Shape", abstract = TRUE)

Circle <- new_class("Circle",
  parent = Shape,
  properties = list(
    radius = positive_numeric
  )
)
```

Here we've included what we've written so far. I used `abstract = TRUE` to tell S7 that we don't ever mean to make a `Shape` object outright. This means the following is now disallowed:
 <!-- with a small change to `Shape` to disallow instantiation of arbitrary `Shape`s. -->

```{r}
#| error: true
Shape()
```

Let's look at some more interesting things we can do with properties.

```{r}
Circle <- new_class("Circle",
  parent = Shape,
  properties = list(
    radius = positive_numeric,
    area = new_property(class_numeric, getter = function(self) pi * self@radius^2, setter = NULL)
  )
)
```

Here, we are defining the area of a `Circle` through a property, but we aren't setting that property explicitly at construction, like we are with `@radius`. Instead, we are creating a [computed property](https://rconsortium.github.io/S7/articles/classes-objects.html#computed-properties), which also happens to be [frozen](https://rconsortium.github.io/S7/articles/classes-objects.html#frozen-properties). What that means is that `@area` is calculated when you call it, and the value itself cannot be set. The `getter` is a function in a property that defines what happens when you call `object@property`. It must always take only one argument, `self`, and returns the value of the property. Here, we can see that `@area`'s getter is defined to grab `@radius` and use that to calculate the area of a circle. The setter defines what happens when you do `obj@property <- value`. In this case, you technically don't need to set setter to `NULL` as that is its default value and would make `@area` an immutable property. I advocate for explicitly setting the setter to `NULL` to make it very clear that `@area` is a read only property.

```{r}
#| error: true
circ <- Circle(10)

circ@area
circ@radius <- 20
circ@area

circ@area <- 0
```

Circles are getting boring; let's add Rectangles and Squares.

```{r}
Rect <- new_class("Rect",
  parent = Shape,
  properties = list(
    width = positive_numeric,
    height = positive_numeric,
    area = new_property(class_numeric, getter = function(self) self@width * self@height, setter = NULL)
  )
)
```

I hope by now the class definition should be pedestrian. The only real difference between `Rect` and `Circle` is that `Rect` has an extra property and a different means to compute area.

```{r}
Rect(10, 6)
Rect(10, 10)
```

And now we can easily extend our shapes to further include `Square`s.

```{r}
Square <- new_class("Square",
  parent = Rect,
  constructor = function(side) {
    force(side)
    new_object(Rect, width = side, height = side)
  }
)
```

Here, we are leveraging the idea that all squares are rectangles to offload most of the heavy work to `Rect`. We are essentially presenting `Square` as an special case `Rect` with well defined behavior. All we need is to define the parent of `Square` as `Rect` (whose parent is `Shape`, whose parent is `S7_object`), and define a short custom constructor. We can set default values of the object in the constructor itself, or could pass a default value in the property definition.

As you can see, the constructor is almost a regular R function--it just ["must always end with a call to `new_object()`"](https://rconsortium.github.io/S7/articles/classes-objects.html#constructors). Another important thing to note is that if you define a custom constructor, ["any subclass will also require a custom constructor"](https://rconsortium.github.io/S7/articles/classes-objects.html#constructors).

```{r}
Square@constructor
Rect@constructor
```

We didn't set a default value for `side` in the constructor, meaning that this won't work:

```{r}
#| error: true
Square()
```

Since `Square`s are `Rect`s, we get all the same nice things `Rect`s do.

```{r}
#| error: true

s <- Square(10)
s
s@area <- 0

Square(-9)
```

We could've overwritten `Rect`'s `@area` to exploit our knowledge of `Square`s:

```{r}
Square <- new_class("Square",
  parent = Rect,
  properties = list(
    area = new_property(class_numeric, getter = function(self) self@width^2, setter = NULL)
  ),
  constructor = function(side) {
    force(side)
    new_object(Rect, width = side, height = side)
  }
)

Square(5)

Square@properties$area$getter
```

Note that we can't use `side` in the definition for `@area` since `side` isn't a property--it's just the name we gave to the only argument we take when creating `Square`s.

There is certainly a lot more to learn about properties and classes, but hopefully this has helped you build enough of an understanding of how these constructs work to be able to reason about them. The website (and of course `{BigNum}`) are great places to see examples and learn more.

## Generics and Methods

[Website](https://rconsortium.github.io/S7/articles/generics-methods.html)

This section is going to be shorter since I think this is best explained on the website and is closer to S3 (which I assume is the OOP most are familiar with) than Properties and Classes.

We have objects, and we can give our objects some data--but we can't do anything with them yet. That's where Generics and Methods step in. A generic defines an idea of a behavior while a method defines an implementation of that behavior for a particular class. [^3] Let's revisit the idea of `Shape`s' area. We initially defined area as a property, but there's no reason why we couldn't make it a function--this may not always be the case.

[^3]: S7 supports multiple dispatch so this is technically not correct--Methods define implementations for an arbitrary set of classes, which could be just one class, but doesn't have to be. Multiple dispatch means that S7 can select a method based on the classes of multiple arguments, not just the first one; i.e. `method(obj1, obj2)` could have different behavior from `method(obj1, obj3)`. S3 only has single dispatch, except for [double dispatch](https://adv-r.hadley.nz/s3.html#double-dispatch)

```{r}
positive_numeric <- new_property(class_numeric,
  validator = function(value) {
    if (value <= 0) "must be greater than 0"
  }
)

Shape <- new_class("Shape",
  properties = list(
    sides = new_property(class_integer,
      validator = function(value) {
        if (value <= 0) "must be greater than 0"
      }
    )
  ),
  abstract = TRUE
)

Circle <- new_class("Circle",
  parent = Shape,
  properties = list(
    radius = positive_numeric
  ),
  constructor = function(radius) {
    force(radius)
    new_object(S7_object(), radius = radius, sides = 1L)
  }
)

Rect <- new_class("Rect",
  parent = Shape,
  properties = list(
    width = positive_numeric,
    height = positive_numeric
  ),
  constructor = function(width, height) {
    force(width)
    force(height)
    new_object(S7_object(), width = width, height = height, sides = 4L)
  }
)

Square <- new_class("Square",
  parent = Rect,
  constructor = function(side) {
    force(side)
    new_object(Rect, width = side, height = side, sides = 4L)
  }
)
```

Let's review what we've changed. First, `Shape`s have recovered `@sides`, with an identical definition to what we saw earlier. To accommodate this, I've written custom constructors for all `Shape`s, and defined `@sides` for each shape there. Additionally, we didn't specify any default values for any shape now, so all arguments are mandatory. Now we wish to get the area of a `Shape`, so we will write a generic, then a method for `Circle`s.

```{r}
Area <- new_generic("Area", "shape")
method(Area, Circle) <- function(shape) {
  pi * shape@radius^2
}
```

The arguments to `new_generic()` are relatively straightforward. We start with the name of the method, which should always be the same as the name you assign the value of `new_generic()` to, just like with Classes. The second argument are the dispatch arguments. S7 uses multiple dispatch, which means that multiple objects can be used to determine the correct method for a given generic. `{BigNum}` uses multiple dispatch (see definition of operators), but for this example we will rely on single dispatch. We pass the title of the argument(s) we wish to dispatch on. We aren't passing the dots here, and [read the website](https://rconsortium.github.io/S7/articles/generics-methods.html#generic-method-compatibility) to see what that implies.

After that comes a method implementation. `method()` is very flexible, and allows you to register methods for a variety of generics: viz. for [an S7 generic, an external generic, an S3 generic, or an S4 generic.](https://rconsortium.github.io/S7/reference/method.html) We specify the generic we are defining a method for, and then the class we are adding behavior to. We implement the actually functionality in a garden variety R function--except it must have the named argument as defined in the generic:

```{r}
#| error: true
method(Area, Circle) <- function(x) {
  pi * x@radius^2
}
```

And now we can calculate the area of `Circle`s.

```{r}
Area(Circle(10))
```

Since we haven't defined methods for any other `Shape`s, this will result in an error.

```{r}
#| error: true
Area(Square(5))
```

With our type hierarchy, we could define a "default" area by creating a method for `Shape`.

```{r}
method(Area, Shape) <- function(shape) {
  "No method defined for this Shape!"
}

Area(Square(5))
method_explain(Area, Square)

method(Area, Rect) <- function(shape) {
  shape@height * shape@width
}

Area(Square(5))
method_explain(Area, Square)
```

Similarly, we could get the number of sides of any `Shape` with minimal effort, due to how we defined `@sides`.

```{r}
number_of_sides <- new_generic("number_of_sides", "x")
method(number_of_sides, Shape) <- function(x) {
  x@sides
}

number_of_sides(Square(5))
number_of_sides(Rect(4, 2))
number_of_sides(Circle(10))

method_explain(number_of_sides, Circle)
```

Let's say that we want to be able to get the perimeter of our `Shape`s--but we don't want to make our own generic. Instead, we want to use `length()` to get the perimeter of a Shape. Doing so is trivial; just remember to use the correct arguments in the method definition.

```{r}
method(length, Circle) <- function(x) {
  2 * pi * x@radius
}

length(Circle(10))

method(length, Rect) <- function(x) {
  2 * (x@height + x@width)
}

length(Square(5))
```

If you need to define behavior for a generic defined in another package, you can use [`new_external_generic()`](https://rconsortium.github.io/S7/reference/new_external_generic.html). Example from the site:

```{r}
median_Shapes <- new_external_generic("stats", "median", "x")
method(median_Shapes, Shape) <- function(x) {
  "You can't take the median of a Shape!"
}

median(Rect(4, 2))
```

Note that this isn't necessary for `base`, `stats`, etc. functions, but would be if you wish to register a method for a generic defined in another package--i.e. an external generic. This example is just to let you see the syntax and usage.

Since S7 is built atop of S3, you could do something like this:

```{r}
median.Shape <- function(x) {
  "Yay S3"
}
median(Rect(4, 2))
```

but I strongly recommend that you use `method()` instead.

There are far more details on the website's [Generics and Methods page](https://rconsortium.github.io/S7/articles/generics-methods.html).

## Conclusion

Equipped with this knowledge of S7, you should hopefully be able to understand how this package and S7 works--or at the very least, know enough to know what to search or learn. I'd suggest reading through the source code of the package next, which I hope is a more interesting (and certainly more fleshed out) example than the `Shape`s we looked at. Additionally, the code in the package is slightly different as it is in a package, unlike the examples shown.
